---
title: "2A - Statistical Inference"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this document we will briefly practice some foundational concepts of 
statistical inference and hypothesis testing.  Follow the instructions 
in the comments of each code chunk.

## Exercise 1 - Running a $t$-test

```{r}
# Download the charity donation dataset from the URL provided
url <- "https://peopleanalytics-regression-book.org/data/charity_donation.csv"

data <- read.csv(url)

set.seed(123)
random_ages <- sample(data,100)


```

```{r}
# Create two vectors capturing the total donations for those who have 
# recently donated and those who have not
recent <- data$total_donations[data$recent_donation == 1]
mean_r <- mean(recent)

not_recent <- data$total_donations[data$recent_donation == 0]
mean_nr <- mean(not_recent)
```

```{r}
# Calculate the difference in the means of these two vectors
# Round this to 2 decimal places.

mean_diff <- round(mean_r - mean_nr,2)

```

```{r}
# Test the hypothesis that recent donors have a different mean donation amount. 
# Ensure that your test is saved as an object with a name of your choice.

test <- t.test(recent,not_recent)

```

```{r}
# EXTENSION:  Test the hypothesis that recent donors donate MORE
# (Hint: seek help on the t.test function to work out how to do this)
help(t.test)


more_test <- t.test(recent,not_recent,
                    alternative = "greater")
```

## Exercise 2 - Interpreting a $t$-test

```{r}
# The results of a t-test are actually a named list.
# You can access specific elements of the list using $ (eg my_test$p.value)
# Return the standard error value for the difference in mean total donations. 
# Round this to 2 decimal places
round(test$stderr, 2)

```

```{r}
# Return the p-value and the 95% confidence interval for the population diff.
# round these to 2 decimal places 
round(test$p.value,2)
round(test$conf.int, 2)



```

Use these values to write an interpretation of your $t$-test, explaining what you have observed in the sample and what this means you can infer about the population.

I am 95% confident that the true difference in total donations for people that donated recently and not recently is between $148.61 and $1414.14.

It is also significant that those that donated recently also donate more than those that did not donate recetnly.


