---
title: "2B - Linear Regression"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this document we will conduct linear regression modeling on a data set relating to graduate salaries in the United States, to try to understand economic factors which influence gradiate salaries,

## Exercise 1 - Running a simple linear regression model

```{r}
# Download the graduates dataset from the URL provided
url <- "https://peopleanalytics-regression-book.org/data/graduates.csv"

data <- read.csv(url)

```

```{r}
# view the first few rows in the data set to acquaint yourself with the data
head(data)
```

```{r}
# Determine how many observations we have in this data set
library("dplyr")
count(data)
```

```{r}
# use the colnames() function to get the names of the columns in the data set
colnames(data)
```

```{r}
# Use the cor() function to determine the correlation between unemployment rate and median salary
cor(data$Unemployment_rate,data$Median_salary)

```

```{r}
# Consider the scale of the Unemployment_rate column.  
# Transform it into a more useful scale
# (Hint: we will want to understand the impact of a percentage point change)
data$Unemployment_rate <- data$Unemployment_rate *100

```


```{r}
# Run a simple linear regression model to estimate the influence of unemployment rate
# on graduate salaries, saving your model using a name of your choice
model <- lm(data = data, formula = Median_salary ~ Unemployment_rate)

```



## Exercise 2 - Interpreting the coefficients

```{r}
# Examine the coefficients of your saved model

summary(model)

model$coefficients

```

```{r}
# Determine the 95% confidence interval for the coefficients

confint(model, level = .95)

```

Write below an interpretation of these coefficients using the estimates and the confidence intervals which you just calculated.

We are 95% confident that the as the unemployment rate increases by one degree the median salary decreases  between 3419 and 1211


```{r}
# EXTENSION:  If you are familiar with using ggplot2, create a plot of unemployment rate and
# median salary and show the estimated linear model using geom_smooth()
library(ggplot2)

plot <- ggplot( data = data, aes(x = Unemployment_rate, y = Median_salary)) + 
                  geom_point()+
                  geom_smooth(method = "lm") +
                  labs(x = "Unemployement Rate (%)",
                       y = "Median Salarry ($)")
  


```

## Exercise 3 - Running a multiple linear regression model

```{r}
# Run a multiple linear regression to determine the influence of both unemployment rate
# and total graduate employees on median salary

model_2 <- lm(data = data, formula = Median_salary ~ Unemployment_rate + Total)

```

```{r}
# View the coefficients and confidence intervals
coefficients(model)


```

Write below an interpretation of the coefficients of this model.


## Exercise 4 - Including categorical input variables

```{r}
# Add discipline as an input variable to the previous model you created in 
# Exercise 3.  
model_3 <- lm(data = data, formula = Median_salary ~ Unemployment_rate + Total + Discipline)

```


```{r}
# Inspect the results and determine which value was used as a reference for 
# discipline in the model
unique(data$Discipline)

summary(model_3)
```


Write below your thoughts on whether any specific disciplines have a significant influence on median salary.

The Disciplines of Computers & Math, Education, Engineering, and Physical Sciences all had P values below .05

```{r}
# EXTENSION:  Change your reference discipline to a discipline of your choice,
# rerun the model and inspect the results.  How have they changed?

```



## Exercise 5 - Determining model fit

```{r}
# Determine the fit of your model from Exercise 4 
summary(model_3)$r.squared

```

```{r}
# Determine if any of the input variables can be removed to form a
# more parsimonious model

model_4 <- lm(data = data, formula = Median_salary ~ Unemployment_rate  + Discipline)

summary(model_4)$r.squared


```

```{r}
# Verify that these is no substantial difference in the fit of your
# more parsimonious model

summary(model_4)

```

```{r}
# Run a summary of your parsimonious model and comment on
# whether you are confident that it is better than a null model
#
#According to the very low p value on my F-statistic, my model is better than a null model

```

## EXTENSION Exercise 6 - Model diagnostics

```{r}
# Determine if the two numerical input variables have a significant correlation
cor.test(data$Unemployment_rate, data$Total)

```

```{r}
# Using the package mctest, use the imcdiag() function to test if collinearity 
# exists in your full model from Exercise 3.  Use the VIF method for this.
library(mctest)

mctest::imcdiag(model_3, method = "VIF")

```


```{r}
# View a density plot of the Median salaries in the data set - what do you observe?

density(data$Median_salary) |> plot(main = "Median Salary Density")

```



```{r}
# use the qqnorm() function to determine how the residuals of your model 
# compare to a normal distribution
# comment on your observations
# consider re-running your model to address any concerns you have

qqnorm(model_3$residuals)

```

