---
title: "03A - Binomial Logistic Regression"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this document we will conduct binomial logistic regression modeling on a data set relating to recruiting data and whether a person is hired or not hired based on test scores, interviews, and gender.  Data fields: 

  - gender: gender
  - sat: SAT test score of applicant
  - gpa: GPA of applicant
  - apttest: Score on an aptitude test given to applicant
  - int1: applicant rating given by one of two line manager interviewers (1-5 where high is positive)
  - int2: applicant rating given by second of two line manager interviewers (1-5 where high is positive)
  - int3: applicant rating given by HR interviewer (1-5 where high is positive)
  - hired: hired (1) or not hired (0)

## Exercise 1 - Running a simple binomial logistic regression model

```{r}
# Download the recruiting dataset from the URL provided
url <- "https://peopleanalytics-regression-book.org/data/recruiting.csv"
recruiting <- read.csv(url)

```

```{r}
# view the first few rows in the data set to acquaint yourself with the data
recruiting
```

```{r}
# What do you think the datatypes are?
# gender: Categorical
# sat: Quantitative
# gpa: Quant
# apttest: Quant
# int1: Quant
# int2: Quant
# int3: Quant
# hired: Boolean
```



```{r}
# Perform any necessary datatype adjustments
colSums(is.na(recruiting)) # no nas in the data

```

```{r}
# get to the know the data a bit.  what is the split of hired/not hired,
# do you notice differences between the interviewers, etc
# Hint: summary(), GGally::ggpairs() are good first steps
summary(recruiting
        )
GGally::ggpairs(recruiting)
```

```{r}
# Run a simple binomial logistic regression model to estimate the influence of SAT scores
# on hire success, saving your model using a name of your choice
model <- glm(hired ~ sat,
             data = recruiting,
             family = "binomial")
model

```


## Exercise 2 - Interpreting the coefficients

```{r}
# Examine the coefficients of your saved model
(coef <- summary(model)$coefficients |> as.data.frame())
coef
```

```{r}
# Take a look at the odds ratio for our independent variable. How do you interpret this?
coef$odds_ratio <- exp(coef[,"Estimate"])
coef

```

```{r}
# Do you think this is a good model?
# looking at the p value, it looks like the SAT is a significant measure if someone is going to be hired or not
```

## Exercise 3 - Running a multiple binomial logistic regression model

```{r}
# What variables look like they might be interesting to include?
# I would use gpa, apttest, and gender to start because I would want to see if HR are actually using those data points. I would also want to see gender alone is a significant predictor to make sure that we are not biased

```

```{r}
# make a new model with gender and apttest

model_2 <- glm(hired ~ gender + apttest,
              data = recruiting, 
              family = "binomial")
model_2
```

```{r}
# View the coefficients and log odds.
(coef_2 <- summary(model_2)$coefficients |> as.data.frame())
coef_2$odds_ratio <- exp(coef_2[,"Estimate"])
coef_2
```


```{r}
# Assess model fit - is this better than the simple model?
DescTools::PseudoR2(model_2,
                    which = c("McFadden", "CoxSnell", "Nagelkerke", "Tjur"))


#predictions
new_data <- data.frame(
  apttest = c(60, 60, 70, 70),
  gender = c("M", "F", "M", "F")
)

predict(model_2, new_data, type = "response")
```

Write a few sentences on the data and the results of your model.  Odds can be difficult to explain - can you try to 
explain to a stakeholder what this means for a candidate with a 60 on their aptitude test vs a candidate with a 70?

It looks like this model is an okay fit to the data. The pseudo R2 numbers are pretty low, but they are not lower than .1.

I would also predict that a candidates with a .60 and .70 would be hired around 13% and 21.5% of the time respectivly with a slightly higher probability for males.
